---
title: 'Class8: Machine Learning'
author: "Fiona Moore"
date: "4/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## K-means clustering 

Let's start with an example of running the kmeans() function.

```{r}
# Generate some example data for clustering
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))

plot(x)
```



Use the kmeans() function setting k to 2 and nstart=20
Inspect/print the results
Q. How many points are in each cluster?- 30
Q. What ‘component’ of your result object details
 - cluster size? - 30
 - cluster assignment/membership?  
 - cluster center? - cluster means
Plot x colored by the kmeans cluster assignment and
 add cluster centers as blue points
 
 
 
```{r}
kmeans(x, 2, 20) 
```

Have to save results as something. 
```{r}
km <- kmeans(x, 2, 20)
```

Cluster size 
```{r}
km$size
```

Cluster  
```{r}
km$cluster
```
Cluster center 
```{r}
km$centers
```
Plot kmeans cluster assignment.  

```{r}
plot(x, col=km$cluster)
points(km$centers, pch=18, col="blue", cex=3)
```

## Hierarchical Clustering 

We must give the hclust() function a distance matrix, not the raw data as an input.

```{r}
# Distance matrix calculation
d <- dist(x)

# Clustering 
hc <- hclust(d)
plot(hc)

```

Finding Clusters
```{r}
plot(hc)
abline(h= 6, col= "red")

```

```{r}
cutree(hc, k = 2)
```

## Clustering 

```{r}
# Step 1. Generate some example data for clustering
x <- rbind(
 matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), # c1
 matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
 matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3
 rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")

# Step 2. Plot the data without clustering
plot(x)

# Step 3. Generate colors for known clusters
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)
```


Use the dist(), hclust(), plot() and cutree() functions to return 2 and 3 clusters.
How does this compare to your known 'col' groups?

```{r}
d <- dist(x)
hc <- hclust(d)
plot(hc)


```

```{r}
gp2 <- cutree(hc, k= 2)
gp2
```

```{r}
gp3 <- cutree(hc, k= 3)
gp3
```

```{r}
plot(x, col= gp3)
```

```{r}
plot(x, col = gp2)
```

Table for "col" groups.
```{r}
table(col)
```

```{r}
table(gp3)
```

```{r}
table(gp2)
```

## PCA 

```{r}
# Example data
mydata <- read.csv("https://tinyurl.com/expression-CSV",
 row.names=1)  

head(mydata)
```

```{r}
nrow(mydata)
```

```{r}
ncol(mydata)
```

```{r}
colnames(mydata)
```

prcomp() expects the samples to be rows and
genes to be columns so we need to first transpose the
matrix with the t() function
```{r}
head(t(mydata))
```

Running PCA on the transpose of our data. 
```{r}
pca <- prcomp(t(mydata), scale=TRUE) 
```

```{r}
# A basic PC1 vs PC2 2-D plot
plot(pca$x[,1], pca$x[,2], xlab= "PC1", ylab= "PC2") 
```

```{r}
# Variance captured per PC
pca.var <- pca$sdev^2 
pca.var.per <- round(pca.var/sum(pca.var)*100, 1) 

pca.var.per
```

PC1 contains most of the variation. 

```{r}
barplot(pca.var.per, main="Scree Plot",
 xlab="Principal Component", ylab="Percent Variation")
```

```{r}
## A vector of colors for wt and ko samples
colvec <- colnames(mydata)
colvec[grep("wt", colvec)] <- "red"
colvec[grep("ko", colvec)] <- "blue"
plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
 xlab=paste0("PC1 (", pca.var.per[1], "%)"),
 ylab=paste0("PC2 (", pca.var.per[2], "%)")) 
```

## PCA of UK Food Data 

```{r}
x <- read.csv("class08.csv", row.names = 1)

head(x)
```

Q1. How many rows and columns are in your new data frame named x? What R functions could you use to answer this questions?

```{r}
nrow(x)
```
```{r}
ncol(x)
```

Plotting the data 
```{r}
barplot(as.matrix(x), beside=T, col=rainbow(nrow(x)))
```

Change the value of "beside." 
```{r}
barplot(as.matrix(x), beside=F, col=rainbow(nrow(x)))
```

Pairwise Plots
```{r}
pairs(x, col=rainbow(10), pch=16)

```
What does it mean if a given point lies on the diagonal for a given plot? 
- if everything was the same then all of the points would lay on the diagonal 
- N. Ireland is the most different from England 

```{r}
# Use the prcomp() PCA function 
pca <- prcomp(t(x))
summary(pca)
```

Q7. Complete the code below to generate a plot of PC1 vs PC2. The second line adds text labels over the data points.

```{r}
# Plot PC1 vs PC2
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x))
```

## Digging Deeper 

```{r}
## Lets focus on PC1 as it accounts for > 90% of variance 
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,1], las=2 )
```



 
 
 
 
