---
title: "Class 13"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set up for Docking 

We will first prepare our HIV-Pr system for drug dicking by making a protein only PDB format file.
```{r}
library(bio3d)
file <- get.pdb("1hsg")
```

```{r}
pdb <- read.pdb(file)
pdb
```

```{r}
prot <- atom.select(pdb, "protein", value = TRUE)
write.pdb(prot, file= "1hsg_protein.pdb")
prot
```

```{r}
lig <- atom.select(pdb, "ligand", value=TRUE)
write.pdb(lig, file="hsg_ligand.pdb")
lig
```

## Inspecting docking results 

In order to visualize the docks and compare to the crystal conformation of the ligand we willprocess the all.pdbqt to a PDB format file that can be loaded into VMD. To do this we will use R and the Bio3D package.

```{r}
library(bio3d)
res <- read.pdb("all.pdbqt", multi=TRUE)
write.pdb(res, "results.pdb")
```

To assess the results quantitatively we will calculate the RMSD (root mean square distance)between each of the docking results and the known crystal structure using the bio3d package.
Back in RStudio read the original ligand with added hydrogens that you produced earlier and use the rmsd() function to compare to your docking results.

```{r}
# res <- read.pdb("all.pdbqt", multi=TRUE)
ori <- read.pdb("1hg_ligand.pdbqt")
rmsd(ori, res)
```

##  Normal Mode Analysis (NMA)

```{r}
pdb <- read.pdb("1hel")
modes <- nma( pdb )
m7 <- mktrj(modes, mode=7, file="mode_7.pdb")
```

Then you can open the resulting mode_7.pdb file in VMD
