---
title: 'Class 7: R functions and packages'
author: "Fiona Moore"
date: "4/23/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Functions revisited

We will source a file from online with our functions from last lecture. 

```{r}
source("http://tinyurl.com/rescale-R")

```

Try out the last day's rescale() function

```{r}
rescale(1:10)
```

Try the rescale2() fucntion that catches string inputs. It has a stop function included.  
```{r}
#rescale2(c(1:10, "string"))
```


#Find missing NA vales in two vectors
Start with a simple example of the larger problem I am trying ti solve

```{r}
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)
```

```{r}
is.na(x)
```
 
```{r}
is.na(y)  
```

Try putting these together with an AND

```{r}
is.na(x) & is.na(y)
```

How many trues are in this answer?
Take the sum() to find how many TRUE values we have and thus how many NAs we had in both x and y

```{r}
sum(is.na(x) & is.na(y))
```

Now I can make this into our first function...

```{r}
both_na <- function(x, y) {
  sum(is.na(x) & is.na(y))
}
```

```{r}
both_na(x, c(NA, 3, NA, 2, NA))
```

Test, test, test

```{r}
x <- c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)
```

```{r}
both_na(x, y2)
```

```{r}
y3 <- c(1, NA, NA, NA, NA, NA, NA)
both_na(x, y3)
```

```{r}
3 == 3
```

```{r}
3 != 2
```
Put a stop() into our fucntion to stop it when the length of x and y are not equal.
Now lets try the both.na2() fucntion on our different length inout vectors.
```{r}
#both_na2(x, y2)
```

```{r}
which(c(F,F, T, F, T))

#which(is.na(c(1, 2, NA, 4)))


```

```{r}
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)

both_na3(x, y)
```


#Another example: intersect function 

Finding common genes in two data sets 

```{r}
df1
```

```{r}
df2
```


Make things simple
```{r}
x <- df1$IDs
y <- df2$IDs

x
y
```

```{r}
intersect(x, y)
```

```{r}
x %in% y
```

```{r}
which(x %in% y)
```

```{r}
x[x %in% y]
```

```{r}
cbind(x[x %in% y], y[y %in% x])
```

Use the R studio shortcut 'Code > Extract Function' to turn our snippet into a working function.
```{r}
gene_intersect <- function(x, y) {
  cbind(x[x %in% y], y[y %in% x])
}
```

```{r}
gene_intersect(df1$IDs, df2$IDs)
```

```{r}
gene_intersect2(df1, df2)
```

```{r}
gene_intersect3(df1, df2)
```

```{r}
merge(df1, df2, by = "IDs")
```

# Grade function
Calculate the average/ mean score, dropping the lowest 

```{r}
x <- c(100, 100, 100, 100, 100, 100, 100, 90)
y <- c(100, 90, 90, 90, 90, 90, 97, 80)
```

```{r}
min(x)
```

```{r}
(sum(x) - min(x)) / (length(x)-1)
```
```{r}
grade <- function(x) {
  (sum(x) - min(x)) / (length(x)-1)
}
```

```{r}
grade(y)
```

 





